---
title: "Analysis of shooting incidents trend in Toronto from 2004 to 2023"
author: 
  - Hyungsoo Park
thanks: "https://github.com/kakaomonk/totonro_shooting_incidents"
date: today
date-format: long
abstract: "The research analyzes the trends of shooting incidents by firearm in Toronto from 2004 to 2023.The used dataset that is from Opendatatoronto provides all shooting incidents that was happened in Toronto throughout the period. In addition, this study also shows the urban violence in Toronto. Fourth sentence."
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
library(dplyr)
library(knitr)
```

# Introduction

Owning firearm by individual has been became a major social issue in Toronto and other urban areas in North America. However, the owning firearm by individual cannot be illegalized since some citizens need the firearm inevitably. The urban violence by firearm in Toronto has been become an object of public concern and being more seriously. Beyond the controversy the study mainly focuses on the trend of firearm shooting incident in Toronto from 2004 to 2023, and there was a significant changes on the trend and the change was enought to make the issue be more controversial.

The firearm violence affect to several social community and it is directly related to the social safety in all communities in Toronto. In addition, Toronto is a large city that is multi-cultural and full of diversities. However, they might cause the hate issues because of the difference between the cultures.

The trends of firearm shooting incident in Toronto throughout the period is depended on many factors from the society, but the purpose of the study is exploring the general trends of firearm shooting incidents in Toronto. Analyzing the incident trends will significantly decrease firearm incidents in Toronto, and derive better laws and rules to the society for the citizens' safety.


Since the shooting incidents by gunfire has been become a social issue





# Data {#sec-data}

The used dataset contains all firearm incidents that caused death and injuries in Toronto from 2004 to 2023, and the dataset is from 'opendatatoronto' 'R' package [@citeOpendatatoronto]. The dataset can be found on 'opendatatoronto' website named with "About Shootings & Firearm Discharges". All used technologies are done by 'R' [@citeR], 'ggplot2' [@citeGGplot2], and 'knitr'[@citeKnitr].

## Data Cleaning
The purpose of the study is to analyze the trends of firearm incidents in Toronto from 2004 to 2023. The preserved columns are related to the date of occurrence and the area code. All columns that are not related to the future analysis were removed. For accurate result, every invalid data that contains uncertain data was omitted and it was not used for the study.

## Data 
Some of our data is of penguins (@fig-bills), from @palmerpenguins.
The count
```{r}
#| echo: false
#| warning: false
#| message: false
#| tbl-cap: "Number of firearm shooting incidents by year"
cleaned_data <- read_csv("/cloud/project/outputs/data/cleaned_shooting_incidents.csv")

incidents_by_year <- cleaned_data |>
  group_by(OCC_YEAR) |>
  count(name = "Incidents")
kable(incidents_by_year)

```


# Results

Our results are summarized in @tbl-modelresults.

#```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

#library(rstanarm)
#
#first_model <-
#  readRDS(file = here::here("outputs/models/first_model.rds"))
#```

#```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

#modelsummary::modelsummary(
#  list(
#    "First model" = first_model
#  ),
#  statistic = "mad",
#  fmt = 2
#)
#```




# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows... 

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

#pp_check(first_model) +
#  theme_classic() +
#  theme(legend.position = "bottom")
#
#posterior_vs_prior(first_model) +
#  theme_minimal() +
#  scale_color_brewer(palette = "Set1") +
#  theme(legend.position = "bottom") +
#  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```



\newpage


# References


